---
title: "Fixing rJava on macOS"
author:
date:
output:
  pdf_document: default
---

```{r setup, cache = F, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# Overview

Java and macOS (Version $\geq 10.12$) do not play well due to some legacy issues. It has an adverse impact on the `rJava` package where a lot of `R` packages, unfortunately, including our beloved `mlr` package, depend on. We are not concerned about "why" but more about "how to fix it". Note that there is no universal solution as it is sensitive to:

* the version of Java JDK
* the version of macOS
* the version of R

Therefore, the following proposed solution might or might not work in your Macbook. Unless the author of `rJava` fixes this issue (which I doubt so) or Oracle finds a solution, this `rJava` issue will be eternal whenever you upgrade any of Java JDK, macOS, and/or R.

# Not-so-universal Solution

## Step 1

Open Terminal and type:

```{r, eval=FALSE}
> /usr/libexec/java_home -V
```

If you have Java JDK 11.02 installed, you should get the following message. Else, go to https://www.oracle.com/technetwork/java/javase/downloads/jdk11-downloads-5066655.html to download it.

```{r, eval = FALSE}
Matching Java Virtual Machines (4):
    11.0.2, x86_64:	"Java SE 11.0.2"	/Library/Java/JavaVirtualMachines/jdk-11.0.2.jdk/Contents/Home
```

## Step 2

To get Java running in R, you call `javareconf` as follows in Terminal

```{r, eval=FALSE}
> sudo R CMD javareconf
```

## Step 3

Let's try our luck by installing and loading the `rJava` package.

1. Download and install `rJava`
2. Load it by calling `library(rJava)`

```{r}
library(rJava)
```


If there is no error message, you are good. Else, go to [Step 4]. 

## Step 4

A side note on the error message above: the `rJava` package fails to load because its "namespace" has gone to the incorrect path to load `libjym.dylib`. To mitigate this, try the following in R console (not Terminal)!

```{r}
options("java.home"="/Library/Java/JavaVirtualMachines/jdk-11.0.2.jdk/Contents/Home/lib")
Sys.setenv(LD_LIBRARY_PATH='$JAVA_HOME/server')
dyn.load('/Library/Java/JavaVirtualMachines/jdk-11.0.2.jdk/Contents/Home/lib/server/libjvm.dylib')
library(rJava)
```

If there is no error, you are good to utilise `rJava`. Otherwise, please seek help from us or Google as soon as possible. Again, keep in mind that these steps might not solve this issue on your Macbook.

