---
title: "R Practice: Chapter 1"
author:
date:
output: 
  beamer_presentation:
    incremental: true
---

# Why R

- Open source and free
- Easy to use
- Many packages
- Generic functions. For example, you can use `plot` for any object

```{r, eval = FALSE}
x <- c(1:4)
plot(x)
```

```{r, eval = FALSE}
mod <- lm( y ~ x1 + x2)
plot(mod)
```

# Why Not R

- Open source and free: no technical support
- Easy to use: not the strongly typed languages
- (Too many) packages
  + Not all packages are equally good
  + Some packages have the similar purposes
- Generic functions
  + CS term: Polymorphic methods with dynamic dispatching
  + Which objects where the `plot` function works?
  + Example: try `plot` on `randomForest` object
  
# R's Frenemies

* Open-Source: Python, C#, C++, Java (Weka), F#...
* Commercial: SAS, MATLAB, IBM SPSS...
* In practice, they can co-exist.
* Be pragmatic.
  + If you can do this in `R`, how can you do this in < >?
  + If you can do this in < >, how can you do this in `R`?
* This course is tailor-made for `R` users
* However, we also accept projects in Python.

# MLR package


* A unified framework of all machine-learning packages in `R`
* Similar to `scikit` from Python
* It follows a more structured machine learning "pipeline"

# R you ready?

* Quick refresher: [R bootcamp](https://sites.google.com/a/rmit.edu.au/intro-to-stats/r-bootcamp) developed by [Dr James Baglin](https://www.rmit.edu.au/contract/staff-contracts/academic-staff/b/baglin-dr-james) (RMIT login identity is required).
* Marathon: 
  + `R` manuals available in `RStudio`
  + `R` Cookbook: Proven Recipes for Data Analysis, Statistics, and Graphics" by Paul Teetor
  + and more...

# R Practices

* Warm-Up examples before the class
* R practices in the last hour of the lecture starting week 2
* No grade on R practices
* See Canvas for more details
* Be patient with us as we are new to Canvas

# Any Question?



