---
title: "Analysis of Trends"
output: html_notebook
---

### 1. The dataset ‘wages’ of TSA package contains monthly values of the average hourly wages (in dollars) for workers in the U.S. apparel and textile products industry for July 1981 through June 1987.

```{r}
rm(list=ls())
# install.packages('TSA')
library(TSA)

data(wages)
# head(wages)
wages
class(wages)
```

#### Display and interpret the time series plot for these data.

#### Use least squares to fit a linear time trend to this time series. Interpret the regression output. Save the standardized residuals from the fit for further analysis.

```{r}
plot(wages)
```
```{r}
# model1 = lm(rwalk~time(rwalk)) # label the model as model1 <- lecture
model1 = lm(wages~time(wages))
summary(model1)

```

Then we return:
Estimates of slope ß^ is β̂ . ~1
= `2.811`

Estimate of intercept is β~0
= `-5.49`

Is the slope statistically significant at 5%?
* Yes because: `p-value: < 2.2e-16`

#### Construct and interpret the time series plot of the standardized residuals from part (b) in a new graphics window.

- Use the `least squares approach` to fit a quadratic time trend to the wages time series. 

```{r}
plot(model1,type='o',ylab='y',, main = "Time series plot for wages")
abline(model1) # add the fitted least squares line from model1
```

- Interpret the regression output. 

- Save the standardized residuals from the fit for further analysis.


- Construct and interpret the time series plot of the standardized residuals from part (d) in a new graphics window.
- Compare the plots from (c) and (e) in terms of the randomness of standardised residuals.

### 2. The dataset “beersales” of TSA package contains monthly U.S. beer sales (in millions of barrels) for the period January 1975 through December 1990.
- Display and interpret the plot the time series plot for these data.
- Now construct a time series plot that uses separate plotting symbols for the various months. Does your interpretation change from that in part (a)?
- Use the least squares approach to fit a seasonal-means trend to this time series. Interpret the regression output. Save the standardized residuals from the fit for further analysis.
- Construct and interpret the time series plot of the standardized residuals from part (c). Be sure to use proper plotting symbols to check on seasonality in the standardized residuals.
- Use least squares to fit a seasonal-means plus quadratic time trend to the beer sales time series. Interpret the regression output. Save the standardized residuals from the fit for further analysis.
- Construct and interpret the time series plot of the standardized residuals from part (e). Again, use proper plotting symbols to check for any remaining seasonality in the residuals.

### 3. Load the monthly commercial landings dataset of US-NMFS from the file “NMFS_Landings.csvPreview the document” into the workspace of R. Do the following tasks over the landings series;

- Display and interpret the time series plot for these data.
- Construct a time series plot that uses separate plotting symbols for the various months. Do you see the seasonal effect better?
- Use least squares to fit a linear time trend to this time series. Interpret the regression output. Save the standardized residuals from the fit for further analysis.
- Use the least squares approach to fit a quadratic time trend to the landings time series. Interpret the regression output. Save the standardized residuals from the fit for further analysis.
- Use the least squares approach to fit a harmonic trend to the landings time series. Interpret the regression output. Save the standardized residuals from the fit for further analysis.
- Conduct residual analyses for all models.
