---
title: "Machine Learning: Chapter 8 - Part I"
subtitle: "R Practice"
author:
date:
output:
  beamer_presentation:
    incremental: no
    theme: "Madrid"
    colortheme: "dolphin"
    fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval =TRUE)
```

# Evaluation methods

* Why do we evaluate?
* How do we implement evaluation measures in R?
* Which measure(s) should we use?

# Evaluation for classification problems in mlr

Key functions you need to know

1. `listMeasure`
2. `performance`
3. `calculateConfusionMatrix`
4. `generateThreshVsPerfData`
5. `plotROCCurves`

Additional functions you need to know

6. `makeCostMeasure` (cost-sensitive classification; to be discussed in Week 10)

# Evaluation for regression problems in mlr

Key functions you need to know

1. `listMeasure`
2. `performance`

Not so many functions

# Important remarks

* Do not assign an object with the measure name. 
* For example, this will cause an error.

```{r, eval = FALSE}
mmce <- "Hello word"
performance(pred, measures = list(mmce))
```

* To solve it, remove the first `mmce`

```{r, eval = FALSE}
rm(mmce)
```

* Multinomial and binary problems share but not all measures.
* Always use `predict.type = "prob"` in defining learners for classification problems!

# Demonstration 

* Binary classification: `Ionosphere` data
* Regression: `Boston Housing` data


# FAQ

