---
title: 'Machine Learning: Chapter 6 - Part II'
subtitle: 'R Practice (OPTIONAL): Naive Bayes'
version: 1
author:
date:
output: 
  pdf_document:
    fig_caption: true
    toc_depth: 4
    number_sections: true
referencecolor: blue
linkcolor: blue
link-citations: yes
bibliography: ref.bib
references:
- id: kelleher
  type: book
  issued:
    - year: 2015
  author:
  - family: Kelleher
    given: John D.
  - family: Namee
    given: Brian Mac
  - family: D'Arcy
    given: Aoife
  publisher: The MIT Press
  title: "Fundamentals of Machine Learning for Predictive Data Analytics"
  subtitle: "Algorithms, Worked Examples, and Case Studies"
- id: hadley
  title: "tidyverse: Easily Install and Load Tidyverse Packages"
  author: 
  - given: Hadley 
    family: Wickham
  issued:
    year: 2016
  url: https://CRAN.R-project.org/package=tidyverse
- id: hadley2
  title: "dplyr: A Grammar of Data Manipulation"
  author: 
  - given: Hadley 
    family: Wickham
  - given: Francois
    family: Romain
  - given: Henry
    family: Lionel
  - given: Müller
    family: Kirill
  issued:
    year: 2017
  url: https://CRAN.R-project.org/package=dplyr
- id: yihui
  author:
  - family: Xie
    given: Yihui
  type: article-journal
  title: "knitr: A General-Purpose Package for Dynamic Report Generation in R"
  issued:
    year: 2016
- id: mlr
  title: "`mlr`: Machine Learning in R"
  author:
  - family: Bischl
    given: Bernd
  - family: Lang
    given: Michel
  - family: Kotthoff
    given: Lars
  - family: Schiffner
    given: Julia
  - family: Richter
    given: Jakob
  - family: Studerus
    given: Erich
  - family: Casalicchio
    given: Giuseppe 
  - family: Jones
    given: Zachary M.
  url: http://jmlr.org/papers/v17/15-066.html
  issued:
  - year: 2016
  volume: 17
  pages: 1-5
  publisher: Journal of Machine Learning Research
- id: ron
  author:
  - family: Kohavi
    given: Ron
  type: article-journal
  title: "Scaling Up the Accuracy of Naive-Bayes Classifiers: a Decision-Tree Hybrid"
  issued:
    year: 1996
  publisher: Proceedings of the Second International Conference on Knowledge Discovery and Data Mining

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

# Objective and Dataset Description

The goal is to classify if a banknote is counterfeit or genuine based on its image information. The dataset is at available at [UCI Machine Learning Repository](http://archive.ics.uci.edu/ml/datasets/banknote+authentication), at the courtesy of Professor [Volker Lohweg](https://www.researchgate.net/profile/Volker_Lohweg) and Dr. Helene Doerksen from University of Applied Sciences, Ostwestfalen-Lippe, Germany. The dataset contains 1372 instances and the following features:

1. Variance of Wavelet Transformed image (continuous) 
2. Skewness of Wavelet Transformed image (continuous) 
3. Curtosis of Wavelet Transformed image (continuous) 
4. Entropy of image (continuous) 
5. Class (1 means counterfeit; 0 means genuine) 

The fifth feature is the target feature. The data were extracted from images taken during the evaluation of an authentication procedure for banknotes. Wavelet transform tools were used to extract features from images. The wavelet transformation is an image-processing technique to analyse and reconstruct multi-resolution images. 

Image processing is beyond the scope of this course. But, we know that the first, second, and third features are statistical properties of the transformed images. Perhaps counterfeit and genuine notes have different variances. The entropy of image (the forth feature) measures the amount of information which must be coded for by a compression method. A low entropy image has very litte contrast, for example, a dark sky. Therefore, it is easier to compress the low entropy images. 


# Questions

We shall use the mean misclassifcation error (mmce) rate to evaluate the model performance.

1. Read and load the data in R. Note that the data has no header. You need to specify `header=FALSE` in `read.csv` function. Process the data if necessary.
2. Create classification task and configure a Naive Bayes learner. 
3. Train the model without resampling.
4. Create the confusion matrix. How many counterfeit notes are misclassified as genuine?
5. Why we cannot apply Laplace smoothing on this dataset?
6. Train the model with sampling. Suppose your learner and task are named as `learner1` and `classif.task`. First, create a three-folded stratified sampling using `makeResampleDesc` function. Let's name it `rdesc`. Then, create a resampled trained model object named `mod` by passing `rdesc` to `resample` function together with `learner1` and `classif.task`.

```{r, eval = FALSE}
rdesc <- makeResampleDesc("CV", iters = 3L, stratify = TRUE)
mod   <- resample(learner1, classif.task, rdesc)
```

7. Obtain the prediction of resampled model as follow

```{r, eval = FALSE}
pred <- getRRPredictions(mod)
```

8. Find the optimal threshold to reduce mmce. Does this threshold also reduce the misclassification rate of counterfeit notes (class = 1)?
