---
title: 'Machine Learning: Chapter 1'
subtitle: 'Solutions to Very Optionl Practice Questions'
version: 1
author:
date:
output: 
  pdf_document:
    fig_caption: true
    toc_depth: 4
    number_sections: true
referencecolor: blue
linkcolor: blue
link-citations: yes
bibliography: ref.bib
references:
- id: kelleher
  type: book
  issued:
    - year: 2015
  author:
  - family: Kelleher
    given: John D.
  - family: Namee
    given: Brian Mac
  - family: D'Arcy
    given: Aoife
  publisher: The MIT Press
  title: "Fundamentals of Machine Learning for Predictive Data Analytics"
  subtitle: "Algorithms, Worked Examples, and Case Studies"
- id: hadley
  title: "tidyverse: Easily Install and Load Tidyverse Packages"
  author: 
  - given: Hadley 
    family: Wickham
  issued:
    year: 2016
  url: https://CRAN.R-project.org/package=tidyverse
- id: hadley2
  title: "dplyr: A Grammar of Data Manipulation"
  author: 
  - given: Hadley 
    family: Wickham
  - given: Francois
    family: Romain
  - given: Henry
    family: Lionel
  - given: Müller
    family: Kirill
  issued:
    year: 2017
  url: https://CRAN.R-project.org/package=dplyr
- id: yihui
  author:
  - family: Xie
    given: Yihui
  type: article-journal
  title: "knitr: A General-Purpose Package for Dynamic Report Generation in R"
  issued:
    year: 2016
- id: mlr
  title: "`mlr`: Machine Learning in R"
  author:
  - family: Bischl
    given: Bernd
  - family: Lang
    given: Michel
  - family: Kotthoff
    given: Lars
  - family: Schiffner
    given: Julia
  - family: Richter
    given: Jakob
  - family: Studerus
    given: Erich
  - family: Casalicchio
    given: Giuseppe 
  - family: Jones
    given: Zachary M.
  url: http://jmlr.org/papers/v17/15-066.html
  issued:
  - year: 2016
  volume: 17
  pages: 1-5
  publisher: Journal of Machine Learning Research

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

# Practice Questions (OPTIONAL)

## Conceptual Questions

1. Which of the following is a **supervised machine learning**?

\color{blue}

> a. Yes. Forecast the electricity consumptions in Melbourne. \color{blue}Yes\color{black}
> b. Yes. Predict the price of Bitcoin next month.
> c. No. Identify unknown customer groups based on their spending habit.
> d. No. Create a recommended YouTube playlist.
> e. Yes. Classifying false news based on Tweets and Facebook newsfeed.

> Items c and d are unsupervised machine learning problems as they do not have target response features.

\color{black}

2. What is **overfitting**?

> \color{blue} Overfitting arises when the prediction model selected by the algorithm is too complex that it fits to the dataset too closely and becomes sensitive to the noise in the data.\color{black}

3. What is **underfitting**?

> \color{blue} Underfitting arises when the prediction model selected by the algorithm is too simplistic to represent the underlying relationship between the target feature and the descriptive features..\color{black}

## Computational Questions

**Question 1**

Download, read and clean Table 1.2 @kelleher[, p. 6].

\color{blue}

Note that the last columns contains `\t`. We need to remove it. There are many approaches to do this. In the first approach, we can apply the `trimws` and `sub` functions on these columns.

\color{black}

```{r, eval = FALSE}
library( tidyverse )
scores2 <- read.csv('Table1-2.csv', skip = 2, stringsAsFactors = FALSE)
scores2$Outcome    <- trimws(sub('\t', '', scores2$Outcome))
scores2$Occupation <- trimws(sub('\t', '', scores2$Occupation))
scores2$Property   <- trimws(sub('\t', '', scores2$Property))
scores2$Type       <- trimws(sub('\t', '', scores2$Type))

```

\color{blue}

In the second approach, we can use the `mutate` function from the `tidyverse` package.

\color{black}

```{r, eval = FALSE}
scores2 <- scores2 %>% 
  mutate( Outcome    = trimws(sub('\t', '', Outcome)) ,
          Occupation = trimws(sub('\t', '', Occupation)),
          Property   = trimws(sub('\t', '', Property)),
          Type       = trimws(sub('\t', '', Type))
  )
```

\color{blue}

The third approach is to apply the `sapply` function with a user-defined function named `subAndTrim`. This approach might be a bit complex but it reduces potential syntax errors. The `apply` family is very useful in implementing the machine learning models. Type `?apply` in the `R` Console for more information and examples how to utilise the `apply` functions.

\color{black}

```{r, eval = FALSE}
subAndTrim <- function( x ){
  
  x <- trimws(sub('\t', '',  x)) 
  x <- factor( x )
  return(x)
}

scores2[, sapply(scores, class) == 'character'] <- sapply( 
  scores2[, sapply(scores, class) == 'character'], FUN = subAndTrim   )    

```

**Question 2**

Apply Rule 1.1 on Table 1.2.

```{r, echo = FALSE}
scores2 <- read.csv('scores.csv')
scores  <- read.csv('scores_short.csv')
```

```{r}
rule1 <- function( data, k = 3){
  
  if(k < 0){
    stop( 'k must be larger than zero.')
  }
  
  data$pred <- ifelse( data$Ratio > k, 'default', 'repay')
  data$pred <- as.factor(data$pred)
  
  return( data[, c('Outcome', 'pred')] )
}

mod1  <- rule1( scores2, k = 3 )
head(mod1)
table(mod1)
```

**Question 3**

Write a function based on Rule 1.2. 

\color{blue}

We have more parameters in Rule 1.2. One approach is to apply the `ifelse` function. Another approach is to utilise the `case_when` function from the `dplyr` package. The expression of the `case_when` function is more readible. 

\color{black}

```{r, message = FALSE, warning = FALSE}
library(dplyr)
rule2 <- function( data, k1, k2, age, occupation ){
  
  data$pred <- case_when(
    (data$Ratio < k1) ~ "repay",
    (data$Ratio > k2) ~ "default",
    (data$Age < age & data$Occupation == occupation) ~ "default",
    TRUE ~ "repay"
  )
  
  data$pred <- as.factor(data$pred)
  
  return( data[, c('Outcome', 'pred')] )
}

```

**Question 4**

Apply this function from previous question on Table 1.1 and Table 1.2. 

```{r}
mod1 <- rule2( scores, k1 = 1.5, k2 = 4, age = 40, occupation = 'industrial')
mod2 <- rule2( scores2, k1 = 1.5, k2 = 4, age = 40, occupation = 'industrial')

head(mod1)
head(mod2)
```


\newpage

# References