---
title: "Machine Learning: Chapter 4 - Part I"
subtitle: "R Practice"
author:
date:
output:
  beamer_presentation:
    incremental: true
---

# mlr Package Framework

* `mlr` package is built upon other machine learning packages
* You cannot run the learner with `mlr` if you have not installed 
* You can install `mlr` and then the learner package later.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(shape)
plot(1, type="n", xlab="", ylab="", xlim=c(0, 1.4), ylim=c(-1, 0.5),
     axes = FALSE)

for (i in 1:3){
  rect(0, 0.5 -0.3*i ,0.75, 0.2 -0.3*i , 
       lwd = 1, col = "skyblue", border = "white")

  Arrows(0.8, 0.4 - 0.3*i, 1 ,0.4 - 0.3*i)
  
  text( x = 0.35, 0.4 - 0.31*i, paste0("Learner package ", i), cex = 1)
}

rect(1.1, 0.25, 1.7, -0.8,  lwd = 1, col = "pink", border = "white")
text( x = 1.3, -0.25 ,"mlr", cex = 2) 
```


# Learner Packages

* How to check if you have the learner package?
* Call `listLearners()`

```{r, warning = FALSE, message = FALSE}
library(mlr)
head( listLearners()[c("class", "package", "installed")])
```

# mlr Flow I

* Configuring learning task

```{r, eval = FALSE}
task  <- makeClassifTask(id = "income", 
                         data = adult, target = "income")
```

* Configuring learner
```{r}
learner <- makeLearner(cl = "classif.rpart", 
                       predict.type = "prob")
```
 
* Training a learner: 

```{r, eval = FALSE}
mod <- train( learner, task)
```

# mlr Flow II

* Making prediction using the trained learner

```{r, eval = FALSE}
pred <- pred(mod, newdata = adult)
```

* Performance evaluation

```{r, eval = FALSE}
performance(pred)
```

* More details, see Chapter 4 Warm-Up Example
* We will cover performance evaluation in depth in Week 9 Lecture

# Resampling in mlr

* Why resampling?
* Example of Cross Validation Resample

```{r, eval = FALSE}
rdesc <- makeResampleDesc("CV", iters = 3)
r     <- resample("classif.rpart", task, rdesc)
```

# Training multiple models

* You can train multiple models (in your project)
* Note we will cover support vector machine (`ksvm`) in Week 12 (if time permits)

```{r, eval = FALSE}
lrns <- list(
  makeLearner(cl = "classif.ksvm", id = "ksvm1" , 
              sigma = 0.2, C = 2),
  "classif.rpart",
  "classif.randomForest"
)
```

# Chapter 4 R Practice

* Goal: Classify sonar signals from 60 descriptive features
* Data set is available at the UCI and Canvas
* For more details, visit Canvas

# Questions?




