---
title: 'Machine Learning: Chapter 4 (Part II)'
subtitle: 'Solutions to Optional R Practice Questions'
version: 1
author:
date:
output: 
  pdf_document:
    fig_caption: true
    toc_depth: 4
    number_sections: true
referencecolor: blue
linkcolor: blue
link-citations: yes
bibliography: ref.bib
references:
- id: kelleher
  type: book
  issued:
    - year: 2015
  author:
  - family: Kelleher
    given: John D.
  - family: Namee
    given: Brian Mac
  - family: D'Arcy
    given: Aoife
  publisher: The MIT Press
  title: "Fundamentals of Machine Learning for Predictive Data Analytics"
  subtitle: "Algorithms, Worked Examples, and Case Studies"
- id: hadley
  title: "tidyverse: Easily Install and Load Tidyverse Packages"
  author: 
  - given: Hadley 
    family: Wickham
  issued:
    year: 2016
  url: https://CRAN.R-project.org/package=tidyverse
- id: hadley2
  title: "dplyr: A Grammar of Data Manipulation"
  author: 
  - given: Hadley 
    family: Wickham
  - given: Francois
    family: Romain
  - given: Henry
    family: Lionel
  - given: Müller
    family: Kirill
  issued:
    year: 2017
  url: https://CRAN.R-project.org/package=dplyr
- id: yihui
  author:
  - family: Xie
    given: Yihui
  type: article-journal
  title: "knitr: A General-Purpose Package for Dynamic Report Generation in R"
  issued:
    year: 2016
- id: mlr
  title: "`mlr`: Machine Learning in R"
  author:
  - family: Bischl
    given: Bernd
  - family: Lang
    given: Michel
  - family: Kotthoff
    given: Lars
  - family: Schiffner
    given: Julia
  - family: Richter
    given: Jakob
  - family: Studerus
    given: Erich
  - family: Casalicchio
    given: Giuseppe 
  - family: Jones
    given: Zachary M.
  url: http://jmlr.org/papers/v17/15-066.html
  issued:
  - year: 2016
  volume: 17
  pages: 1-5
  publisher: Journal of Machine Learning Research
- id: ron
  author:
  - family: Kohavi
    given: Ron
  type: article-journal
  title: "Scaling Up the Accuracy of Naive-Bayes Classifiers: a Decision-Tree Hybrid"
  issued:
    year: 1996
  publisher: Proceedings of the Second International Conference on Knowledge Discovery and Data Mining

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

# Conceptual Questions


1. Besides `NA`, list 3 other possible values to represent missing values in a given data set.

\color{blue}

> Excessive white space, "-99999" for strictly positive features, "NULL" 

\color{black}

2. Give an example where missing values contain information.

\color{blue}

> $X:=$ Your favorite AFL (Australian Football League) team. If the individual does not like football, $x = NA$. Therefore, any missing value likely indicates the individual does not like football.

\color{black}

3. Would scaling the continuous features change the correlation matrix? The scaling of a variable $x$ is defined as:

$$x \longrightarrow \frac{x-\mu_x}{\sigma_x}$$

\color{blue}

> Correlation is invariant to linear scaling. However, note that correlation will change in non-linear transformation.

\color{black}

4. Give an example where two variables are non-linearly correlated but have a zero correlation.

\color{blue}

> Let $X$ be a standard normal distribution of zero mean and a variance of 1. That is $\mathbb{E}(X)=0$ and $\mathbb{E}(X^2)=1$. Consider $Y = X^2$. $X$ and $Y$ are quadratically related, but they have a zero correlation because:

$$\text{Cov}(X, Y) = \text{Cov}(X, X^2) = \mathbb{E}(X \times X^2) - \mathbb{E}(X^2) \mathbb{E}(X) = 0 - 1\times0 = 0$$

> Note that $\mathbb{E}(X^3) = 0$ as the skewness is zero for a standard normal random variable.

\color{black}

\newpage

# Computational Questions

## Pre-process and explore [Adult Census Income Data](https://archive.ics.uci.edu/ml/machine-learning-databases/adult/) by @ron.

\color{blue}

See the sample project for Phase 1.

\color{black}

## (Optional) Replicate Table 3.3(b)

Let's subset the data with the categorical features.

```{r, warning=FALSE, message = FALSE}
library(mlr)
library(tidyverse)
fraud                         <- read.csv('MotorInsuranceFraudClaimABTFull.csv')
categorical_features          <- fraud[,  sapply( fraud, is.factor )]
quality_report_categorical    <- summarizeColumns( categorical_features ) %>%
  select(-type, -mean, -disp, -median, -mad , -min, -max)    # remove the unwanted columns

```
Before we append the first and the second modes to the quality report. Let's do some experiment on one categorical feature, say `Injury.Type`. We can create the table of `Injury.Type` and assign it as `x`

```{r}
x <- table( categorical_features$Injury.Type )
x
```

To obtain the frequency of the first mode, we simply apply `max(x)` 

```{r}
max(x)
```

To obtain the first mode, we do the following:

```{r}
names(x)[x == max(x)]
```

Now let's utilise these tricks in the `sapply` function

```{r}
quality_report_categorical$modeFreq   <- sapply( categorical_features, function(x){ max(table(x)) }  )
quality_report_categorical$mode       <- sapply( categorical_features, 
                                               function(x){ y <- table(x); names(y)[y == max(y)]    }  )
quality_report_categorical$modePercent<-quality_report_categorical$modeFreq/nrow(fraud)*100
```

How about the second mode? We can subset `x` by excluding `Broken Limb`

```{r}
z <- x[ x != max(x) ]
```

Hence, the second modes and its frequency can be compute as:

```{r}

names(z)[z == max(z)]
max( z )
```

And, we can define a function named `gimmeSecondMode` and `gimmeSecondModeFreq` to return the second mode and its frequency as follow. (Note there is probably a more elegant solution). 
```{r}
gimmeSecondMode <- function( feature){
  
  tab  <- table( feature )
  
  tab  <- tab[ tab != max(tab) ]
  
  if( length(tab) == 0){
    mode <- NA
  }else{
    mode <- names(tab)[tab == max(tab)]
  }
  

  return( mode )
}

gimmeSecondModeFreq <- function( feature){
  tab  <- table( feature )
  tab  <- tab[ tab != max(tab) ]
  
  if( length(tab) == 0){
    max <- NA
  }else{
    max <- max(tab)
  }
  

  return( max )
}
```


```{r}
quality_report_categorical$modeFreq2    <- sapply( categorical_features, gimmeSecondModeFreq )
quality_report_categorical$mode2        <- sapply( categorical_features, gimmeSecondMode     )
quality_report_categorical$mode2Percent <-quality_report_categorical$modeFreq2/nrow(fraud)*100
quality_report_categorical
```


# References

