---
title: 'Machine Learning: Chapter 4 - Part I'
subtitle: 'R Practice Questions (OPTIONAL)'
version: 1
author:
date:
output: 
  pdf_document:
    fig_caption: true
    toc_depth: 4
    number_sections: true
referencecolor: blue
linkcolor: blue
link-citations: yes
bibliography: ref.bib
references:
- id: kelleher
  type: book
  issued:
    - year: 2015
  author:
  - family: Kelleher
    given: John D.
  - family: Namee
    given: Brian Mac
  - family: D'Arcy
    given: Aoife
  publisher: The MIT Press
  title: "Fundamentals of Machine Learning for Predictive Data Analytics"
  subtitle: "Algorithms, Worked Examples, and Case Studies"
- id: hadley
  title: "tidyverse: Easily Install and Load Tidyverse Packages"
  author: 
  - given: Hadley 
    family: Wickham
  issued:
    year: 2016
  url: https://CRAN.R-project.org/package=tidyverse
- id: hadley2
  title: "dplyr: A Grammar of Data Manipulation"
  author: 
  - given: Hadley 
    family: Wickham
  - given: Francois
    family: Romain
  - given: Henry
    family: Lionel
  - given: Müller
    family: Kirill
  issued:
    year: 2017
  url: https://CRAN.R-project.org/package=dplyr
- id: yihui
  author:
  - family: Xie
    given: Yihui
  type: article-journal
  title: "knitr: A General-Purpose Package for Dynamic Report Generation in R"
  issued:
    year: 2016
- id: mlr
  title: "`mlr`: Machine Learning in R"
  author:
  - family: Bischl
    given: Bernd
  - family: Lang
    given: Michel
  - family: Kotthoff
    given: Lars
  - family: Schiffner
    given: Julia
  - family: Richter
    given: Jakob
  - family: Studerus
    given: Erich
  - family: Casalicchio
    given: Giuseppe 
  - family: Jones
    given: Zachary M.
  url: http://jmlr.org/papers/v17/15-066.html
  issued:
  - year: 2016
  volume: 17
  pages: 1-5
  publisher: Journal of Machine Learning Research
- id: ron
  author:
  - family: Kohavi
    given: Ron
  type: article-journal
  title: "Scaling Up the Accuracy of Naive-Bayes Classifiers: a Decision-Tree Hybrid"
  issued:
    year: 1996
  publisher: Proceedings of the Second International Conference on Knowledge Discovery and Data Mining

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

# Pre-requisites

Ensure you have installed the packages below

```{r, message = FALSE, warning = FALSE}
library(rpart)
library(mlr)
library(rpart.plot)
```

# Sonar Data

This was the data set used by @gorman1988analysis in their study to classify sonar signals. The data set is available at [UCI](http://archive.ics.uci.edu/ml/datasets/connectionist+bench+(sonar,+mines+vs.+rocks)). The alternative is to scrap this data from the `mlbench` package [@mlbench]:

```{r}
library(mlbench)
data(Sonar)
```

The dataset contains:

* One target feature which is `Class`. `Class` has two labels: `M` and `R`. `M` represents the sonar signals off a **metal** cyclinder at various angles. `R` represents the sonar signals off a **rock** cyclinder under the similar conditions.
* 60 descriptive features. Each feature is a pattern recorded from the signals in the range of 0 to 1. The features are named as `V01`, `V02`, ... and `V60`. Each number represents the energy within a particular frequency band, integrated over a certain period of time.


# Questions

Build a decision-tree model to classify the sonar signal from the Sonar data with `mlr` [@mlr].

1. Create a classification task
2. Configuring a `rpart` (decision-tree) learner with `predict.type = "prob"`
3. Training the learner
4. \color{blue}(Optional) \color{black} Visualise the trained decision-tree model
5. Making prediction
6. Performance evaluation with the mean misclassification error (mmce) rate
7. \color{blue}(Optional) \color{black} Determine the optimal threshold
8. Train the model by 2-fold Cross Validation resampling
9. Train the model by 3-fold Cross Validation resampling. Does one more fold improve the model accuracy?

\newpage

# References
