---
title: 'Machine Learning: Chapter 6 - Part II'
subtitle: 'Warm-Up Examples: Naive Bayes'
version: 1
author:
date:
output: 
  pdf_document:
    fig_caption: true
    toc_depth: 4
    number_sections: true
    highlight: zenburn
referencecolor: blue
linkcolor: blue
link-citations: yes
bibliography: ref.bib
references:
- id: kelleher
  type: book
  issued:
    - year: 2015
  author:
  - family: Kelleher
    given: John D.
  - family: Namee
    given: Brian Mac
  - family: D'Arcy
    given: Aoife
  publisher: The MIT Press
  title: "Fundamentals of Machine Learning for Predictive Data Analytics"
  subtitle: "Algorithms, Worked Examples, and Case Studies"
- id: hadley
  title: "tidyverse: Easily Install and Load Tidyverse Packages"
  author: 
  - given: Hadley 
    family: Wickham
  issued:
    year: 2016
  url: https://CRAN.R-project.org/package=tidyverse
- id: hadley2
  title: "dplyr: A Grammar of Data Manipulation"
  author: 
  - given: Hadley 
    family: Wickham
  - given: Francois
    family: Romain
  - given: Henry
    family: Lionel
  - given: Müller
    family: Kirill
  issued:
    year: 2017
  url: https://CRAN.R-project.org/package=dplyr
- id: yihui
  author:
  - family: Xie
    given: Yihui
  type: article-journal
  title: "knitr: A General-Purpose Package for Dynamic Report Generation in R"
  issued:
    year: 2016
- id: mlr
  title: "`mlr`: Machine Learning in R"
  author:
  - family: Bischl
    given: Bernd
  - family: Lang
    given: Michel
  - family: Kotthoff
    given: Lars
  - family: Schiffner
    given: Julia
  - family: Richter
    given: Jakob
  - family: Studerus
    given: Erich
  - family: Casalicchio
    given: Giuseppe 
  - family: Jones
    given: Zachary M.
  url: http://jmlr.org/papers/v17/15-066.html
  issued:
  - year: 2016
  volume: 17
  pages: 1-5
  publisher: Journal of Machine Learning Research
- id: ron
  author:
  - family: Kohavi
    given: Ron
  type: article-journal
  title: "Scaling Up the Accuracy of Naive-Bayes Classifiers: a Decision-Tree Hybrid"
  issued:
    year: 1996
  publisher: Proceedings of the Second International Conference on Knowledge Discovery and Data Mining

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

# Overview

We shall use a dataset from a loan application fraud detection [@kelleher, Table 6.2] to show how to implement Naive Bayes classifier with `mlr` [@mlr]. The dataset can be downloaded at http://machinelearningbook.com/teaching-materials/. Also, we shall show why we need Laplace smoothing parameter, $k$ and how to determine the optimal value of $k$.

# Preliminaries

We need the `mlr` [@mlr] and `tidyverse` packages [@hadley] for machine learning and data wrangling.

```{r, warning = FALSE, message = FALSE}
library(mlr)
library(tidyverse)

```

`mlr` calls the `naiveBayes` function from the `e1071` package [@e1071]. Therefore, we need to check if this package has been installed by calling `require(e1071)` or

```{r, message = FALSE, warning=FALSE}
listLearners() %>% 
  filter(class == "classif.naiveBayes") %>% 
  select(class, name, package, installed)
```

Let's read and process the data. Note that the following preprocessing is \color{red} not elegant \color{black}. \color{blue}Can you think of a better way? \color{black}

```{r}
data    <- read.csv('Table6-2.csv', skip = 1)
data$ID <- NULL
data <- sapply(data, function(x){ trimws(sub("\t", '', x)) }) 
data <- data.frame(data)
```

# Naive Bayes 

Let's define the classification task and Naive Bayes learner as follow. Notice that it is preferable to set the predict type to "prob" in `makeLearner` so that we can refine prediction by adjusting the threshold probability.

```{r}
classif.task <- makeClassifTask(id = "fraud", 
                                data = data, 
                                target = "Fraud",
                                positive = 'true')

learner1 <- makeLearner(cl = "classif.naiveBayes", 
                        predict.type = "prob")
```

We train model to make prediction with the full (training) data. Resampling with cross validation is left as an exercise (keep in mind that we only have 20 observations). With a threshold probability of 50 %, we reach `mmce` of 20 % but have misclassified 3 true frauds, which is shown by `calculateConfusionMatrix(pred1)`. This is because of a relatively lower proportion of frauds. This is coined as "imbalanced class problem" where at least one class of the target feature has a very low percentage in the dataset. We shall learn how to deal with this issue in the upcoming weeks.

```{r}
mod1  <- train( learner1, classif.task)
pred1 <- predict(mod1, newdata  = data)

performance(pred1)
calculateConfusionMatrix(pred1)
pred1$threshold

```

# Laplace Smoothing

## A Motivating Example

Consider a *prospective* loan application (i.e. no target feature) below:

* `history = 'paid'`
* `CoApplicant = 'guarantor'`
* `Accommodation = 'false'`

The objective is predict if this application is fradulent. We can create one datum in `R` by defining the following `data.frame` object. Note that the names of the descriptive features or columns must match those of the training data. Otherwise, the `predict` function will not work. Let's name the `data.frame` object `test_data`. 

```{r}
test_data <- data.frame(History = 'paid', 
                        CoApplicant = 'guarantor', 
                        Accommodation = 'false')
```

By applying `filter`, we can find that the training dataset do not contain the same instance of `test_data`.

```{r}
data %>% filter(History == 'paid', 
                CoApplicant == 'guarantor', 
                Accommodation == 'false')

```

Therefore, when we make prediction on `test_data`, it would result a probability value of fraud close to zero. This is why we require Laplace smoothing.

```{r}
predict(mod1, newdata  = test_data)
```

## Fine-Tuning Laplace Parameters

We can call `getParamSet` to check the list of parameters (or arguments) available in `classif.naiveBayes`. In this example, we have only one - `laplace` which is essentially the Laplace smoothing parameter. The result also shows that the values of `laplace` range from `0` to `inf` (i.e. $k \geq 0$).

```{r}
getParamSet("classif.naiveBayes")
```

Let's configure the discrete tune parameter setting where $k = 0, 0.1, 2, 3, 5, 10, 20$ in the chunk below. An alternative would be `makeNumericParam`.

```{r}
ps <- makeParamSet(
  
  makeDiscreteParam('laplace', values = c(0, 0.1, 0.5, 2, 3, 5, 10, 20))
  
)
print(ps)
```

In addition, we need to define the control and 2-folded stratified resampling. Two folds are used for illustration and also because we have a very small sample size. For result reproducibility, specify a random seed.

```{r}
ctrl  <- makeTuneControlGrid()
rdesc <- makeResampleDesc("CV", iters = 2L, stratify = TRUE)
set.seed(123)
```

Let's run the tune-tuning process.

```{r}
res <- tuneParams("classif.naiveBayes", 
                   task = classif.task, 
                   resampling = rdesc,
                   par.set = ps, 
                   control = ctrl, 
                   show.info = FALSE)

```

The result is that the mean test error is 30 %. As expected, it is higher than the error generated by a trained model without fine-tuning. The optiomal Laplace parameter is 3. 

```{r}
res
mmce$minimize
res$x
res$y

```

To vindicate the result, we need to assess how error rate changes with a range of Laplace parameter values by assigning a `generateHyperParsEffectData(res)` object named `psdata`. To visualise it, we can apply `plotHyperParsEffect` on `psdata`. The plot shows that the error rate is minimised when Laplace parameter is larger or equal to 3. Since we only have a range of discrete values, the error rate curve looks a bit kinked.

```{r}
generateHyperParsEffectData(res)
psdata <- generateHyperParsEffectData(res)

plotHyperParsEffect(psdata, x = "laplace", 
                    y = "mmce.test.mean",
                    plot.type = "line")
```

## Fusing Optimal Laplace Parameter

There are two fusing methods. The first is to call `setHyperPars` as follow by incorporating the optimal (Laplace) parameter value via `par.vals = res$x`. The advantage of `setHyperPars` is that the learner no longer needs to run tune-fining process. The downside is that it can only result in a response prediction, not probability.

```{r, eval = FALSE}
tunedLearner0 <- setHyperPars( makeLearner("classif.naiveBayes"), 
                               par.vals = res$x)

```

The second method is to apply `makeTuneWrapper` with tune-funing configuration. It basically repeats tune-fining process. Therefore, it is viable to configure a tune wrapper bypassing `tuneParams` function. Likewise, we can train and predict using the tune wrapper

```{r, cache = TRUE}
tunedLearner1 <- makeTuneWrapper(learner1, rdesc, mmce, ps, ctrl, show.info = FALSE)

tunedMod1  <- train(tunedLearner1, classif.task)
tunedPred1 <- predict(tunedMod1, newdata  = data)

```

Despite the same fine-tuning configurations, the tuned wrapper results in a lower error rate than `res` (from the previous section). \color{red} Why?\color{black} Due to sampling, there are more misclassification of fraudulent applications. The tuned wrapper makes only one correct guess.

```{r}
performance(tunedPred1)
res$x

calculateConfusionMatrix(tunedPred1)

```

However, the Laplace-tuned wrapper is able to predict `test_data`. The probability of a fradulent application is no longer close to zero but lower than the threshold probability of 50 %. Consequently, it is still classified as non-fradulent. 

```{r}
tunedPred1$threshold
predict(tunedMod1, newdata  = test_data)

```

## Adjusting Probability Threshold

By running the codes below, we can improve the accuracy by adjusting the optimal probability threshold to 30 %. \color{red}How do you interpret a lower threshold level? \color{black}

```{r}
d <- generateThreshVsPerfData(tunedPred1, measures = list(mmce))

plotThreshVsPerf(d) + 
  geom_hline(yintercept = 0.3, color = 'red', linetype = 'dashed')

```

By setting the threshold of 30 % on the tuned predictions, we can obtain a new set of predictions which yields the comparable confusion matrix generated by the base Naive Bayes classifier.

```{r}
tunedPred3 <- setThreshold(tunedPred1, 0.3)
calculateConfusionMatrix(tunedPred3 )

```

# References